Curvature Enhanced Shape With and Adapted Laplacian Operator For Hybrid Quad/Triangle Meshes

### 2013-01-14 trunk/documents/thesis/TQLBO-paper/Paper-Siggraph/paper.tex
msg

The LBO has desirable properties: the LBO points out to the direction
of the minimal surface area, minimizing the energy of the total surface
curvature $S$ at equation (\ref{eq:total_cuarvature_integral}).

@@

The method allows shape magnification and exaggeration using the laplacian
smoothing operator in the reverse direction, generating a parameterized
family of shapes from a set of vertices which represent a coarse sketch
of the desired model. Our approach can be mixed with traditional or
uniform subdivision surfaces methods, is iterative and converges towards
a continuous and smooth version of the original model.

Our method allows the use of soft constraints weighting the effect
of smoothing at each vertex based on a normalized weight, which are
assigned to the control vertices of the original mesh or. The weights
of the new vertices resulting from the subdivisions are calculated
by interpolation, allowing modifying the behavior of the method on
specific regions of the original model.

Our work presents an extension of the Laplace Beltrami operator for
hybrid quad/triangle meshes, allows using mixed arbitrary types of
mesh representation, exploiting the basic geometrical relationships
facilitating and ensuring convergence of the algorithm and similar
shapes consistent with the original shape against the other methods.

@@

Given a mesh $M=\left(V,Q,T\right)$, with vertices $V$, quads $Q$,
triangles $T$.

The area of $1$-ring neighborhood ($N_{1}$) with shared face quad
or triangle to vertex $v_{i}$ in $M$ is.

@@

Applying the mean average area according to \cite{Xiong2011} of all
possible triangulations for each quad to $A\left(Q_{N_{1}\left(v_{i}\right)}\right)$
as show in figure \ref{fig:quad_xiong}.

@@

Thus triangles and quads configurations of the 1-ring neighborhood
faces adjacent to $v_{i}$ can be simplified to five cases shown in
figure \ref{fig:LBO-basic-5-TQ}.

Then according to equation (\ref{eq:discrete_mean_curvature_normal}),
(\ref{eq:def_LBO}), and five simple cases defined in figure \ref{fig:LBO-basic-5-TQ}
the TQLBO (Triangle-Quad LBO) of $v_{i}$ is.

@@

\subsection{Curvature Enhancing\label{sub:Curvature-Enhancing}}

@@

\caption{\label{fig:Subdivision-Cups} Family of cups generated with our method
from coarse model to enhancing the curvature obtained from Catmull-Clark
Subdivision and the use of constraints over coarse model with weight
vertex group in red.}

@@

The curvature enhancing uses the inverse direction of the curvature
flow to move the vertices in the portions of the mesh with most curvature.
In this equation we use a diffusion process:

@@

To solve the equation above we use implicit integration and a normalized
version of TQLBO matrix.

@@

The vertices $V^{t+1}$are enhanced along their inverse curvature
normal directions by solving the linear system: $Ax=b$, where $A=I-\left|\lambda dt\right|W_{p}L$,
$L$ is the Normalized TQLBO defined in the equation (\ref{eq:TQLBO-Normalized_Matrix}),
$x=V^{\prime}$ are the smoothing vertices, $b=V^{t}$ are the actual
vertices positions, $W_{p}$ is a diagonal matrix with weight vertex
group, and $\lambda dt$ is the enhance factor that support negative
and positive values, negative for enhancing positive for smoothing.

The curvature cannot be calculted at the border of the meshes that
are not closed, for that reason we use the scale-dependent operator
proposed by \cite{Desbrun1999}.

Our method was designed for use with weighted vertex groups which
specify the final curvature enhancement on the solution, these weights
vary between $0$ and $1$ with a value of $0$ makes no changes and
with values 1 applies a maximal change. The weights modify influence
zones where the Laplacian is applied as shown in the equation \ref{eq:Lineal_System_with_wp}.
Families of shapes that are generated may change substantially with
the weights of specific control points.

The models volume increases as the lambda is larger and negative,
this can be go against by a simple method of volume preservation.
In \cite{Desbrun1999} presents a simple method to resize the mesh
however the model suffers large displacements when $\lambda<-1.0$
or after multiple iterations. We propose the following solution: If
$v_{i}^{t+1}$ is a mesh vertex of $V^{t+1}$ in the $t+1$ iteration,
we define $\overline{v}$ as:

@@

$\overline{v}$ is the center of the mesh, $vol_{ini}$ is an initial
volume, and $vol_{t+1}$ is the volume at the iteration $t+1$, then
we define a scale factor.

@@

The new vertices positions are:

@@

\caption{\label{fig:Catmull_Clark}Left: Original Model, in green color model
with Catmull-Clark Subdivision. Models with Laplacian smoothing: $\lambda=60.0$,
$\lambda_{e}=12.0$ and $\lambda=90.0$, $\lambda_{e}=80.0$. Models
first filter with Laplacian smoothing $\lambda=60.0$, $\lambda_{e}=12.0$
and before applied curvature enhancing: $\lambda=-200.0$, $\lambda_{e}=12.0$
and $\lambda=-40.0$, $\lambda_{e}=-100.0$.}
\end{figure*}

We designed a new brush that allows the enhancement of the curvatures
of a polygon mesh in real time. Our brush works well with the stroke
method named \textsl{Drag Dot}  which allows to preview the changes
in the model before you release the mouse button or pen tablet, also
allows to move the mouse over the model to fit exactly where you want
to perform the enhancement of the curvature.

Brushes that perform similar work as inflate can create distortions
in the mesh and can also produce self-intersections in the mesh, as
these brushes only move the vertices along the normal and do not take
into account the global information. Whereas our method looks for
the best way to make inflation while preserving the global curvature
retaining the several shape and sharp features of the model.

Our method simplifies the work required for the enhancement, which
would be to use some different brushes for inflating and some other
to soften and styling. With our enhanced brush in one step can be
performed.

For the real-time brush it necessary that the Laplacian matrix is
constructed with the vertices that are within the radius  of the sphere
defined by the user, which reduces the size of the matrix to be processed,
the center of this sphere depends on the place where the user clicks
on the canvas and three-dimensional mesh place where the click is
projected.

Furthermore special handling is required for the vertices at the boundary
which have neighbors that are not within the radius of the brush.
These vertices must be marked as boundary and the curvature will not
be calculated for them, but must be present in the matrix so that
every vertex has their neighbors within the selection, this change
allows the results to be much smoother on the border. The Laplacian
matrix for sculpt mode as a matrix equation.

@@

Where $v_{j}\in N\left(v_{i}\right)$, $u$ is the center of sphere
of radius $r$. The matrixes should remove rows and columns of vertices
index that are not within the radius.

@@

The Catmull-Clark subdivision transformation is used to smooth a surface
as the limit of a sequence of subdivision steps\cite{Stam1998}. This
process is governed by properties of B-splines curve from multivariate
spline theory\cite{Loop1987}. This method does a recursive subdivision
transformation that refines the model into a linear interpolation
that approximates a smooth surface. The smoothness of the model is
automatically guaranteed \cite{DeRose1998}.

Catmull-Clark subdivision surface methods generate smooth and continuous
models from a coarse model and produce results quickly due to the
simplicity of implementation, but with these methods is not easy to
make changes to the global curvature of the model. If we use Catmull-Clark
subdivision surfaces and curvature enhancement for modeling from coarse
models with few vertices can generate families of shapes changing
a single parameter, this would allow an artist to choose the model
from a similar set of options that best meets his/her needs without
having to change each one of the control vertices.

Our method allows the use of vertex weight paint over the control
points. The weights can be applied to a coarse model, then to perform
a Catmull-Clark subdivision where weights are interpolated, producing
weights with smooth changes in the zones of influence, so the obtained
curvature is much softer at those areas as shown in figure \ref{fig:Subdivision-Cups}.c.

In the equation \ref{eq:Lineal_System_with_wp}, $W_{p}$ is a diagonal
matrix with weights corresponding to each vertex. Weights at each
vertex produce a different solution for this reason the matrix is
placed in the diffusion equation, as families that are generated may
change substantially with weighted of specific control points.

@@

In this section we describe the results of our curvature enhance method
with the extension of the Laplace Beltrami operator for hybrid quad/triangle
meshes with several example models (see figures \ref{fig:Spectrum},
\ref{fig:Subdivision-Cups}, \ref{fig:Catmull_Clark}, \ref{fig:TQLBO_test},
\ref{fig:camello_enhanced}, \ref{fig:Sculpt_Brush}, \ref{fig:Performance-sculpt},
\ref{fig:(a)Monkey}, \ref{fig:Animated_Camell}). We test the curvature
enhance with TQLBO method on a PC with AMD Quad-Core Processor @ 2.40
GHz and 8 GB RAM.

Figure \ref{fig:TQLBO_test} show the results when applying Laplace
Beltrami Operator TQLBO of equation (\ref{eq:TQLBO_Simple_Matrix})
in a model with a simple subdivision. In column (c) Laplacian smoothing
was applied to the model consists of only quads. In column (d) the
model was converted to triangles and then Laplacian smoothing was
applied. In column (e) the model is converted randomly some quads
into triangles and then Laplacian smoothing is applied, showing similar
results to those composed only of triangles or squares.

@@

Methods using Catmull-Clark subdivision surface and the enhancement
allows to modify the curvature that is obtained with the process of
subdivision as shown in figure \eqref{fig:Subdivision-Cups} this
test used a coarse model of a cup, after the model was performed subdivision
then was performed Laplacian smoothing and enhancement by modifying
the parameters $\lambda$ y $\lambda_{e}$ corresponding to the lambda
for rings and edges respectively. In figure \eqref{fig:Subdivision-Cups}.c,
\eqref{fig:Subdivision-Cups}.d shows also the use weight vertex groups
over coarse models with subdivision surfaces that allowed to generate
the weights for the new interpolated vertices, these new weights were
used for the enhancement obtained the 6 cups that are to the right
of the figure \eqref{fig:Subdivision-Cups}.d.

Laplacian smoothing applied with simple subdivision may produce similar
results to those obtained with Catmull-Clark whose models are on average
equal triangles as shown in figure \ref{fig:Catmull_Clark} green
model and that obtained with Laplacian smoothing $\lambda=60.0$,
$\lambda_{e}=12.0$, but also can modify the curvature obtained after
applying Catmull-Clark as shown in the three columns to the right
of the figure \ref{fig:Catmull_Clark}.

Figure \ref{fig:camello_enhanced} show the generation of different
versions of a camel according to parameter lambda. In the top row
you can see results of do curvature enhance over all model, as the
lambda becomes larger and negative in this model it inflates the convex
parts as shown in figure \ref{fig:Spectrum}. The larger $\lambda$
the larger enhancements on the model features. The bottom row of figure
\ref{fig:camello_enhanced} shows the use of weighted vertex groups,
which allows to specify which areas will be enhanced. On the left
the enhancement of the legs of the camel is shown which produces enhancement
of organic aspect, notice that the border is not distorted and smooth
in the union.

@@

\caption{\label{fig:camello_enhanced}Top row: Original camel model in left.
Curvature enhancing with $\lambda=-30.0$, $\lambda=-100.0$, $\lambda=-400.0$.
Bottom row: Curvature enhancing with weight vertex group, $\lambda=-50.0$
and 2 iterations at legs, $\lambda=-200.0$ and 1 iteration in head
and neck.}
\end{figure}

Our method for enhancement of the silhouette features is predictable
and invariant under isometric transformations as those present in
some animations (see Figure \ref{fig:Animated_Camell}). This animation
shows some poses of a camel during a walk, enhancement is performed
in the neck and legs as shown in the bottom left part of the figure
\ref{fig:Animated_Camell}. Local modifications produced by pose interpolation
or animation rigging do not significantly affect the result as with
camel legs in each sample pose different flexion of the leg joints
enhancement camel keeps flesh-like shapes in the pattern produced
original by the artist. This is due to the diffusion process which
is subjected to the mesh so that small local changes are treated without
significantly affecting global solution. Our method is invariant of
rotation, this since depends exclusively on the normal field of the
mesh, wish is invariant under global rotations.

Figure \ref{fig:Sculpt_Brush} shows the use of curvature enhancement
brush for sculpting in real time. One pass was used with the brush
as shown in the figure with the blue and red radius. In figure \ref{fig:Sculpt_Brush}.b
camel shows how the leg self-intersecting and looks like two bubbles
stuck, the same happens to the fingers on the bottom of the figure.
Using silhouette features enhancement in figure \ref{fig:Sculpt_Brush}.c
we get better results sd the are retained shape of the silhouette
and the details of the fingers and leg. Similar results can be obtained
by an artist however it would take several steps and require the use
of several brushes. With curvature enhancement it only takes one step.
This new method can easily enhance organic features like muscles during
the sculpting process. In figure \ref{fig:Performance-sculpt} we
show the performance of curvature enhancement brush, in this experiment
we use three models with 12K, 40K and 164K vertices, this models were
sculpted with curvature enhance brush in each step the brush sphere
contain a variable number of vertices for processing. The processing
time for 800 vertices in the camel paw (40k model) only take 0.1 seconds,
for 2600 vertices in leg and neck (model 40k) take 0.5 seconds, these
times are good, because usually an artist sculpts a model for parts,
and each part is represented by an average of about 3000 vertices
in the models we use.

Tests with the Laplacian operator equation (\ref{eq:TQLBO_Simple_Matrix})
and its normalized version equation (\ref{eq:TQLBO-Normalized_Matrix}),
produce similar results if the triangles that compose the mesh are
the same size on average. The normalized version is more stable and
predictable because it is not divided by the area of the ring which
may be very small and cause problems during floating point calculations
as shown in figure \ref{fig:(a)Monkey}.c bottom row. The enhancement
of curvature of the model with the normalized Laplacian operator has
a more regular behavior. The model can be deformed in TQLBO normalized
version with large lambdas ($\lambda>400$) while it intersects itself
it produces no peaks. Figure \ref{fig:(a)Monkey} shows different
results due to the area of the triangles in the model. Triangles with
larger area have smaller enhancement (figure \ref{fig:(a)Monkey}.c
skull), and smaller triangles have larger enhancements (figure \ref{fig:(a)Monkey}.c
chin).

@@

\caption{\label{fig:Sculpt_Brush}Top row: (a) Leg Camel, (b) Inflate brush
for leg into blue circle, (c) Enhance curvature brush for leg into
red circle. Bottom row: (a) Hand, (b) Inflate brush for fingers into
blue circle, (c) Enhance curvature brush for fingers in red circle.}
\end{figure}

@@

\caption{\label{fig:Performance-sculpt}Performance of our dynamic curvature
enhance brush in terms of vertices sculpted per second. Three models
with 12K, 40K, 164K vertices used for sculpting in real time.}
\end{figure}

@@

\caption{\label{fig:Animated_Camell}Our method is pose insensitive. The enhanced
for the different poses are similar in terms of shape. Top row: Original
walk cycle camel model. Bottom row: Curvature enhancing with weight
vertex group, $\lambda=-400$ and $2$ iterations. }
\end{figure*}

@@

Our method was implemented as a modifier for modeling and brush for
sculpting on the Blender software \cite{blender} with C and C++.
Working with Blender allowed us to test the method interactively against
others as Catmull-Clark, weight vertex groups and sculpting system
in Blender.

To improve the performance, we worked with Blender mesh struct visiting
each triangle or quad and adding its corresponding index in a list
that stores the sum of the Laplacian weights of the ring, in this
way we only had to visit two times the list of faces of the mesh list
and two times the list of the edges if the mesh was not closed. The
brush sculpting mode required to create a list that maps the index
of the selected vertices to a list from 1 to N where N is the number
of vertices selected and the number of rows in the linear system to
solve. This drastically reduced the calculations enabling real-time
processing. In the construction of our Laplacian matrix several index
were locked at vertices having faces areas or lengths edges with value
zero that can cause spikes and bad results.

Under this conditions matrix at the equation\ref{eq:TQLBO_Simple_Matrix}
is sparse since the number of neighbors per vertex corresponding to
the number of data per row is small compared to the total number of
vertices in the mesh. To solve the linear system equation \ref{eq:Lineal_System_with_wp}
was used OpenNL which is a a library for solving sparse linear system.

@@

This work presented an extension of the Laplace Beltrami operator
for hybrid quad/triangle meshes that can be used in production environments
without modification that provides results similar to those obtained
by working only with triangles or quads.

We introduced a new way to perform silhouettes enhancement in a mesh
for modeling or sculpting in a few steps which allows the modification
of the curvature of a model while preserving its overall shape. We
introduce a new method of modeling and show some of its possible uses.
This method behaves in predictable ways which facilitate the learning
process, and works well with isometric transformations opening the
possibility of introducing it on the process of animation.

We show that this tool could work in the early stages in where coarse
models are used, allowing to modify the curvature generated by the
Catmull-Clark subdivision surfaces, avoiding the edition of the vertices
with the change of a few parameters.

As future work we analyze theoretically the relationship between the
Catmull-Clark subdivision surfaces and Laplacian smoothing because
in some cases can produce very similar results. But the subdivision
surface is a fast method thereby reducing computation times for calculate
curvature in the mesh.

>>>>

%i.e.= id est (esto es)
The LBO has desirable properties: the LBO points out to the direction
of the minimal surface area, minimizing the energy of the total surface
curvature $S$ at equation (\ref{eq:total_cuarvature_integral}).

@@

This method exaggerates a shape using a Laplacian smoothing operator in the reverse direction, i.e., the new shape is a modified version in which those areas with larger curvature are magnified. The operator amounts to a generator of a set of models which conserves the basic silhouette of the original shape. In addition, the presented approach can be easily mixed with traditional or uniform subdivision of surfaces. This method is based on an original extension of the Laplace Beltrami operator for hybrid quad/triangle meshes, mixing arbitrary types of meshes, exploiting the basic geometrical relationships and ensuring algorithm convergence.

@@

Given a mesh $M=\left(V,Q,T\right)$, with vertices $V$, quads $Q$,
triangles $T$. The area of $1$-ring neighborhood $A\left(v_{i}\right)$ with shared face quad $A\left(Q_{N_{1}\left(v_{i}\right)}\right)$
or triangle $A\left(T_{N_{1}\left(v_{i}\right)}\right)$ to vertex $v_{i}$ in $M$   is.

@@

Applying the mean average area, according to \cite{Xiong2011}, from all
possible triangulations, as show in figure \ref{fig:quad_xiong}, the area for each quad $A\left(Q_{N_{1}\left(v_{i}\right)}\right)$ is

@@

Triangle and quad configurations of the 1-ring neighborhood
faces, adjacent to $v_{i}$, can be simplified to five cases, as shown in
figure \ref{fig:LBO-basic-5-TQ}.

According to equation (\ref{eq:discrete_mean_curvature_normal}),
(\ref{eq:def_LBO}), and five simple cases defined in figure \ref{fig:LBO-basic-5-TQ}
the TQLBO (Triangle-Quad LBO) of $v_{i}$ is.

@@

\subsection{The Curvature Enhancement\label{sub:Curvature-Enhancing}}

@@

\caption{\label{fig:Subdivision-Cups} Family of cups generated with our method,
from a coarse model: the curvature, obtained from the Catmull-Clark
Subdivision, is enhanced. Soft constraints, over the coarse model, is drawn in red.}

@@

The curvature is enhanced by using the reverse direction of the curvature
flow, moving the vertices towards those mesh portions with larger curvature.
A standard diffusion process is applied:

@@

To solve this equation, implicit integration is used as well as a normalized
version of TQLBO matrix.

@@

The vertices $V^{t+1}$ are enhanced, along their reverse curvature direction, by solving the linear system: $Ax=b$, where $A=I-\left|\lambda dt\right|W_{p}L$,
$L$ is the Normalized TQLBO defined in the equation (\ref{eq:TQLBO-Normalized_Matrix}),
$x=V^{\prime}$ are the smoothing vertices, $b=V^{t}$ are the actual
vertices positions, $W_{p}$ is a diagonal matrix with weight vertex
group, and $\lambda dt$ is the enhancement factor that supports negative
and positive values: negative for enhancement and positive for smoothing.

The method was devised to use with weighted vertex groups, which
specify the final curvature enhancement of the solution, meaning $0$ as no changes and 1 when a maximal change is applied. The weights modify the influence
zones, where the Laplacian is applied, as shown in equation \ref{eq:Lineal_System_with_wp}. Interestingly, the generated family of shapes may change substantially with
the weights of specific control points.

The model volume increases as the lambda is larger and negative,
this can be counteracted with a simple volume preservation. 
However, the mesh may suffer large displacements when $\lambda<-1.0$
or after multiple iterations. A simple volume conservation algorithm is: If
$v_{i}^{t+1}$ is a mesh vertex of $V^{t+1}$ in the $t+1$ iteration,
we define $\overline{v}$ as:

@@

$\overline{v}$ is the mesh center, $vol_{ini}$ is an initial
volume, and $vol_{t+1}$ is the volume at the iteration $t+1$, then
the scale factor.

@@

allows to move the new vertices to:

@@

\caption{\label{fig:Catmull_Clark}Left: Original Model,  color model
with Catmull-Clark Subdivision in green. Models with Laplacian smoothing: $\lambda=60.0$,
$\lambda_{e}=12.0$ and $\lambda=90.0$, $\lambda_{e}=80.0$. Models with a first Laplacian filtering $\lambda=60.0$, $\lambda_{e}=12.0$
and before applying curvature enhancement: $\lambda=-200.0$, $\lambda_{e}=12.0$
and $\lambda=-40.0$, $\lambda_{e}=-100.0$.}
\end{figure*}

A new sculpting brush is herein proposed and aims to inflate the shape, magnifying the shape curvatures
of a polygon mesh in real time. This brush works properly with the stroke
method  \textsl{Drag Dot}, allowing to pre-visualize the model changes before the mouse is released. Also,
it allows to move the mouse along the model to match the curvature zone which is supposed to be enhanced.

Brushes that perform a similar inflation can introduce mesh distortions or produce mesh self-intersections, provided these brushes only move the vertices along the normal without any global information. In contrast, the present method searches for a proper inflation while preserving the global curvature, retaining the original shape and main model features. In addition, this method simplifies the work required for the enhancement since it needs not different brushes for inflating, softening or styling. The enhanced brush can make all these operation in a single step. 
Real-time brushes require the Laplacian matrix is constructed with the vertices that are within the sphere radius defined by the user, reducing the matrix to be processed,
the center of this sphere depending on the place where the user clicks on the canvas and the three-dimensional mesh placed where the click is
projected. Special handling is required for the boundary vertices with neighbors that are not within the brush radius: these vertices are marked as boundary and the curvature is not there calculated, but they must be included in the matrix so that every vertex has their corresponding neighbors within the selection. The sculpting Laplacian matrix reads as.

@@

Where $v_{j}\in N\left(v_{i}\right)$, $u$ is the sphere center
of radius $r$. The matrices should remove rows and columns of vertices that are not within the radius.

@@

The Catmull-Clark subdivision transformation is used to smooth a surface,
as the limit of a sequence of subdivision steps\cite{Stam1998}. This
process is governed by a B-spline curve \cite{Loop1987}, performing a recursive subdivision transformation that refines the model into a linear interpolation
that approximates a smooth surface. The model smoothness is
automatically guaranteed \cite{DeRose1998}.

Catmull-Clark subdivision surface methods generate smooth and continuous
models from a coarse model and produce quick results because of the
simplicity of implementation. Nevertheless, changes to the global curvature are hardly implantable. The Catmull-Clark
subdivision surfaces together with curvature enhancement can easily generate families of shapes by changing
a single parameter, allowing to handle a model with very few vertices. In practice, this would allow an artist to choose a model
from a similar set of options that  would meet his/her needs without having to change each of the control vertices. Likewise, the presented method allows the use of vertex weight paint over the control points. The weights can be applied to a coarse model, followed by  a Catmull-Clark subdivision where weights are interpolated, producing
weights with smooth changes in the influence zones,  as shown in figure \ref{fig:Subdivision-Cups}.c.

In equation \ref{eq:Lineal_System_with_wp}, $W_{p}$ is a diagonal
matrix with weights corresponding to each vertex. Weights at each
vertex produce a different solution so that the matrix must be
placed in the diffusion equation since families that are generated may
change substantially with weighted of specific control points.

@@

The results of the curvature enhancement method
with the extension of the Laplace Beltrami operator for hybrid quad/triangle
meshes with several example models (see figures \ref{fig:Spectrum},
\ref{fig:Subdivision-Cups}, \ref{fig:Catmull_Clark}, \ref{fig:TQLBO_test},
\ref{fig:camello_enhanced}, \ref{fig:Sculpt_Brush}, \ref{fig:Performance-sculpt},
\ref{fig:(a)Monkey}, \ref{fig:Animated_Camell}). The curvature
enhancement was assessed with TQLBO method on a PC with AMD Quad-Core Processor @ 2.40
GHz and 8 GB RAM.

Figure \ref{fig:TQLBO_test} shows the results when applying the Laplace
Beltrami Operator TQLBO of equation (\ref{eq:TQLBO_Simple_Matrix})
in a model with a simple subdivision. In column (c) the Laplacian smoothing
was applied to a model consisting of only quads. In column (d) the
model was converted to triangles and then the Laplacian smoothing was
applied. In column (e) the model was randomly converted  from some quads
into triangles and then the Laplacian smoothing is applied, showing similar
results to those meshes composed only of triangles or quads.

@@

Methods using the Catmull-Clark subdivision surface and the enhancement
allows to modify the curvature that is obtained with the process of
subdivision, as shown in figure \eqref{fig:Subdivision-Cups}. This
test used a coarse cup model, in which the subdivision was performed, followed by a Laplacian smoothing and enhancement. In figure \eqref{fig:Subdivision-Cups}.c,
\eqref{fig:Subdivision-Cups}.d shows also the use of weight vertex groups over coarse models, with subdivision surfaces that allowed to generate
the weights for the new interpolated vertices. These new weights were
used for the enhancement obtained on the 6 cups that are at the right
of the figure \eqref{fig:Subdivision-Cups}.d.

Laplacian smoothing applied with simple subdivision may produce similar results to those obtained with Catmull-Clark, whose models are in average equal triangles, as shown in figure \ref{fig:Catmull_Clark}, green model. The one obtained with the Laplacian smoothing is shown in panel (c), (d)  and those curvature modified versions are in (d) and (f). As can be observed, different versions of the original sketch can be obtained by parameterizing a single model value, a great advantage of the presented method.  Figure \ref{fig:camello_enhanced} shows the generation of different versions of a camel according to the $\lambda$ parameter. In the top row,  it is shown the curvature enhancement results, as $\lambda$ becomes larger and negative, the resultant shape is observed as if the model would inflate the more convex parts, as shown in figure \ref{fig:Spectrum}. The larger the $\lambda$ parameter the larger the model feature enhancements. The bottom row of figure \ref{fig:camello_enhanced} shows the use of weighted vertex groups, specifying which areas will be enhanced. On the left, the enhancement of the camel legs produces an organic aspect, notice that the border is not distorted and smooth.

@@

\caption{\label{fig:camello_enhanced}Top row: Original camel model at left.
Curvature enhancement with $\lambda=-30.0$, $\lambda=-100.0$, $\lambda=-400.0$.
Bottom row: Curvature enhancement with weight vertex group, $\lambda=-50.0$
and 2 iterations for the legs, $\lambda=-200.0$ and 1 iteration for the head
and neck.}
\end{figure}

The enhancement of the silhouette features is predictable
and invariant under isometric transformations, as those classically used in
some animations (see Figure \ref{fig:Animated_Camell}). In this figure, the animation
shows some camel poses during a walk, the enhancement is performed
at the neck and legs, as shown in the bottom left camel in figure
\ref{fig:Animated_Camell}. Local modifications produced by the pose interpolation
or animation rigging practically do not affect the result. In spite of at any pose of the camel legs there is a clear difference, the enhancement method allows a  flesh-like shape in the original pattern produced by the artist. This is due to the mesh restricted diffusion process so that small local changes are treated without affecting the global solution. The method therefore is rotation invariant since it depends exclusively on the normal mesh field, which is rotation invariant.

Figure \ref{fig:Sculpt_Brush} shows the use of a curvature enhancement
brush for sculpting in real time. One pass was used with the brush,
as shown in the figure, with the blue and red radius. In figure \ref{fig:Sculpt_Brush}.b
the camel foot shows the inflation intersection that looks like two bubbles, a similar pattern to what is observed to the fingers on the bottom of the same figure. The silhouette enhancement is observed in figure \ref{fig:Sculpt_Brush}.c since the main shape is retained together with its finger and leg details. Similar results can be obtained by a user, however it would take several steps and require the use
of several brushes, while the curvature enhancement took a single step. Likewise, this new method can easily enhance organic features like muscles during the sculpting process. In figure \ref{fig:Performance-sculpt} the curvature enhancement brush performance is illustrated, in this experiment three models with 12K, 40K and 164K vertices, were used. These models were
sculpted with the curvature enhance brush, at each step the brush sphere
containing a variable number of vertices for processing. The processing
time for 800 vertices in the camel paw (40k model) only took 0.1 seconds,
for 2600 vertices in the leg and neck (model 40k) took 0.5 seconds, these
times are suitable in real applications since an artist sculpts a model for parts,
and each part is represented by an average of about 3000 vertices.

Tests with the Laplacian operator  (equation \ref{eq:TQLBO_Simple_Matrix})
and its normalized version (equation \ref{eq:TQLBO-Normalized_Matrix}),
produce similar results if the triangles that compose the mesh are about
the same size. The normalized version is more stable and
predictable because it is not divided by the area of the ring which
may be very small and causes numeric problems, as shown in figure \ref{fig:(a)Monkey}.c bottom row. The curvature enhancement of the model with the normalized Laplacian operator results in a more regular pattern. The model can be deformed with a TQLBO normalized
version with large $\lambda$ ($\lambda>400$) while intersecting itself with no peaks. Figure \ref{fig:(a)Monkey} shows different
results due to the triangle areas in the model. Triangles with
larger area have smaller enhancement (figure \ref{fig:(a)Monkey}.c
skull), and smaller triangles have larger enhancements (figure \ref{fig:(a)Monkey}.c
chin).

@@

\caption{\label{fig:Sculpt_Brush}Top row: (a) Leg Camel, (b) Inflate brush
for leg into blue circle, (c) Enhance curvature brush for leg into
red circle. Bottom row: (a) Hand, (b) Inflate brush for fingers into
blue circle, (c) Curvature enhancement brush for fingers in red circle.}
\end{figure}

@@

\caption{\label{fig:Performance-sculpt}Performance of our dynamic curvature
enhancement brush in terms of the sculpted vertices per second. Three models
with 12K, 40K, 164K vertices used for sculpting in real time.}
\end{figure}

@@

\caption{\label{fig:Animated_Camell}The method is pose insensitive. The enhancement
for the different poses are similar in terms of shape. Top row: Original
walk cycle camel model. Bottom row: Curvature enhancement with weight
vertex group, $\lambda=-400$ and $2$ iterations. }
\end{figure*}

@@

The method was implemented as a modifier for modeling and brush for
sculpting, on the Blender software \cite{blender} in C and C++.
Working with the Blender allowed to test the method interactively against
others, as Catmull-Clark, weight vertex groups and sculpting system
in Blender.

To improve the performance, it was worked with the Blender mesh struct, visiting
each triangle or quad and storing its corresponding index and the sum of the Laplacian weights of the ring in a list so that only two visits were required for the list of mesh faces and two times the edge list, if the mesh was not closed. This drastically reduced calculations, enabling real-time
processing. In the construction of the Laplacian matrix, several index were locked at vertices having face areas or edge lengths with null value that could cause spikes and bad results.

Under these conditions, the matrix at equation\ref{eq:TQLBO_Simple_Matrix}
is sparse since the number of neighbors per vertex, corresponding to
the number of data per row, is smaller compared to the total number of
vertices in the mesh. To solve the linear system equation \ref{eq:Lineal_System_with_wp}
was used OpenNL which is a a library for solving sparse linear system.

@@

This work presented an extension of the Laplace Beltrami operator
for hybrid quad/triangle meshes that can be used in production environments
and provides results similar to those obtained
by working only with triangles or quads. This paper has introduced a new way to change silhouettes in a mesh for modeling or sculpting in a few steps by means of the curvature model modification while preserving its overall shape. In addition,  a new modeling method has also been presented some possible applications have been illustrated. The method works properly with isometric transformations, opening the
possibility of introducing it on the process of animation.

We show that this tool may work in early modeling stages, case in which coarse
models are used, allowing to modify the curvature generated by the
Catmull-Clark subdivision surfaces, and thereby avoiding edition of the vertices
with change of a single parameters.

Future work includes the analysis of theoretical relationships between the
Catmull-Clark subdivision surfaces and the Laplacian smoothing since they can produce very similar results.
